---
title: "INFO 526 Project 1 Write-up"
subtitle: ""
author: 
  - name: "Pixel Prophets"
    affiliations:
      - name: "College of Information Science, University of Arizona"
description: "Write-up for performing data visualization on the Great American Coffee Taste Test dataset"
format:
   html:
    code-tools: true
    code-overflow: wrap
    embed-resources: true
editor: visual
execute:
  warning: false
  echo: false
---


## Introduction

The dataset The Great American Coffee Taste Test, created by James Hoffmann and coffee company Cometeer, consists of survey data with answers provided by YouTube viewers during the “Great American Coffee Taste Test” event held by the creators mentioned earlier. There are in total 57 variables and 4042 observations in this dataset. From the 57 variables, 13 variables are stored as numeric class and 44 variables are stored as character class. Out of the total 57 variables, the variables that will be used for our project are “age”, “gender”, “strength”, “total_spend”, “caffeine”, and “employment_status”.

The variable “age” describes the age group to which the surveyee belongs. The variable “gender” describes the gender with which the surveyee identifies. The variable “strength” describes the strength of coffee that the surveyee prefers, which is also known as coffee bitterness level. The variable “total_spend” describes the surveyee's monthly spending on coffee (in dollars). The variable “caffeine” describes the surveyee's caffeine level preference. The variable “employment_status” describes the employment status of the surveyee. Below are the unique values for each of these six variables before the data cleaning process.

```{r loadDataset, echo=TRUE}
# load the original dataset
coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')
```

```{r unique_values_age, echo=TRUE}
# For variable "age"
unique(coffee_survey$age)
```

```{r unique_values_gender, echo=TRUE}
# For variable "gender"
unique(coffee_survey$gender)
```

```{r unique_values_strength, echo=TRUE}
# For variable "strength"
unique(coffee_survey$strength)
```

```{r unique_values_total_spend, echo=TRUE}
# For variable "total_spend"
unique(coffee_survey$total_spend)
```

```{r unique_values_caffeine, echo=TRUE}
# For variable "caffeine"
unique(coffee_survey$caffeine)
```

```{r unique_values_employment_status, echo=TRUE}
# For variable "employment_status"
unique(coffee_survey$employment_status)
```

## Question 1: Explore the coffee bitterness preference across age groups and genders

The first question is: Is there any relationship between the variable age groups, gender and coffee bitterness ? We have used the variables “age” , “gender” and “strength” to address this question. The reason we are interested in answering this question is that it helps us understand the strength preferences of coffee drinkers across different age groups and genders.This can help companies market various coffee products based on the demographics and also innovate new products catering to specific age groups and genders.

We used two violin plots to address this question. One plot maps out the strength preferences across genders and another one maps out the strength preferences across age groups. The reason for choosing the violin plot is that firstly, it shows the full distribution of data across each category which is crucial if we want to understand the spread and density of preferences; secondly, it also provides insights into how the preferences are concentrated and where the distribution is wider or thinner. It allows for easy comparison of the distributions across categories within the same plot. The violin plots make it very clear when coffee preferences are skewed (leaning more towards a certain level of bitterness) or multimodal (having more than one peak and representing multiple common preferences). Coffee bitterness preferences may not follow a symmetric distribution (e.g., more people might prefer stronger coffee than weaker coffee). Violin plots can display the distribution of such a non-symmetric dataset.

Before plotting, we need to do some data cleaning on the original dataset. Data cleaning includes two steps: creating a subset containing the three variables then removing rows with any NA values from this subset, and modifying some variables into values suitable for the visualization.

First, create a subset and remove NA rows.

```{r q1_subset_removeNAs, echo=TRUE}
q1_subset = coffee_survey[, c("age", "gender", "strength")]
q1_subset_naremoved = na.omit(q1_subset)
```

In the original dataset, specifically for the gender column there are six different unique values, including “non-binary” and “prefer not to say”. To simplify, we merged the gender column into three categories – “male”, “female” and “others”. In addition, we modified the "strength" column to use numerical values from 1 to 5 to replace the coffee bitterness levels from weak to very strong.

```{r q1_modified, echo=TRUE}
# modify "gender" column
q1_subset_naremoved$gender[q1_subset_naremoved$gender == 'Other (please specify)'] = 'Others'
q1_subset_naremoved$gender[q1_subset_naremoved$gender == 'Non-binary'] = 'Others'
q1_subset_naremoved$gender[q1_subset_naremoved$gender == 'Prefer not to say'] = 'Others'

# modify "strength" column
q1_subset_naremoved$strength[q1_subset_naremoved$strength == 'Weak'] = 1
q1_subset_naremoved$strength[q1_subset_naremoved$strength == 'Somewhat light'] = 2
q1_subset_naremoved$strength[q1_subset_naremoved$strength == 'Medium'] = 3
q1_subset_naremoved$strength[q1_subset_naremoved$strength == 'Somewhat strong'] = 4
q1_subset_naremoved$strength[q1_subset_naremoved$strength == 'Very strong'] = 5
q1_subset_naremoved$strength = as.numeric(q1_subset_naremoved$strength)

# the clean dataset for Question 1
q1_clean = q1_subset_naremoved
```

Now after these cleaning steps are completed, we proceed to create the violin plot using the code below.

```{r q1_plot_1, echo=TRUE}
library(ggplot2)

# Violin plot for Age Group vs Caffeine Level Bitterness
ggplot(q1_clean, aes(x = age, y = strength, fill = age)) +
  geom_violin(trim = FALSE, scale = "width", width = 0.8) +  
  scale_y_continuous(limits = c(1, 5), breaks = c(1, 2, 3, 4, 5), 
                     labels = c("Weak", "Somewhat Light", "Medium", "Somewhat Strong", "Very Strong")) +
  labs(title = "Coffee Bitterness Preferences by Age Group",
       x = "Age Group",
       y = "Coffee Bitterness Level") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(size = 12))
```

```{r q1_plot_2, echo=TRUE}
# Violin plot for Gender vs Caffeine Level Bitterness 
ggplot(q1_clean, aes(x = gender, y = strength, fill = gender)) +
  geom_violin(trim = FALSE, scale = "width", width = 0.8) +  
  scale_y_continuous(limits = c(1, 5), breaks = c(1, 2, 3, 4, 5), 
                     labels = c("Weak", "Somewhat Light", "Medium", "Somewhat Strong", "Very Strong")) +
  labs(title = "Coffee Bitterness Preferences by Gender",
       x = "Gender",
       y = "Coffee Bitterness Level") +
  scale_fill_manual(values = c("Male" = "lightblue", "Female" = "pink", "Others" = "lightyellow")) +  # Set custom colors
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),  # Center x-axis labels
        plot.title = element_text(hjust = 0.5),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(size = 12))
```

From the first violin plot, the most obvious observation is that the youngest age group (surveyees who are less than 18 years old) has the most evenly distributed preferences for coffee bitterness, compared to all other age groups. Among these levels, they show slightly more preference for the “medium” level. As for the other age groups, they mostly prefer “medium” or “somewhat strong” levels for coffee bitterness. Especially for age groups 25-34 and 35-44 years old, even more data points are clustered in these two levels. This reflects how young people may not be used to the bitterness of coffee yet like the adults are. Young to middle-aged adults prefer stronger coffee among adults due to their stress from careers and work.

From the second violin plot, we can observe that surveyees mostly prefer “medium” or “somewhat strong” for coffee bitterness across genders. Females have the highest number of “medium”, while males have the highest number of “somewhat strong” preferences. Others have comparatively more widespread preferences, with the highest number of preferences for “somewhat strong”. This reflects the possible slight differences across gender groups for coffee bitterness preferences.

## Question 2: Explore the distribution of caffeine level preference and monthly spending on coffee across age groups and employment status

The second question is: Is there any relationship among the variables age groups, monthly spending on coffee, preferred caffeine level and employment status? The variables we need to answer this question are “age”, “total_spend”, “caffeine” and “employment_status”. The reason why we are interested in this question is that we are able to detect trends and extract insights for coffee consumption preferences as well as spending habits for coffee from different age.groups and employment status. This can help coffee companies to accurately target consumers with different coffee products to maximize profits.

The plot that we choose to answer this question is the faceted count plot. This visualization choice is the best to address this question because based on the original survey dataset, the survey questions contributed to these four variables all required categorical inputs. There are fixed options for the surveyees to choose from. This means that there are overlapping data points for each combination of values of these four variables. In order to visualize the counts distribution regarding the combination of these four variables within the same plot, the faceted count plot can be implemented to efficiently achieve this goal.

Before making the visualization, we need to do data cleaning on the original dataset. Data cleaning includes two steps (similar to the steps for Question 1): creating a subset containing the four variables then removing rows with any NA values from this subset, and modifying some variables into values suitable for the visualization.

First, create a subset and remove NA rows.

```{r q2_subset_removeNAs, echo=TRUE}
q2_subset = coffee_survey[, c("age", "total_spend", "caffeine", "employment_status")]
q2_subset_naremoved = na.omit(q2_subset)
```

Then, convert the categorical values of the variable “total_spend” into numeric values, with the mean value of each spending range (for the \>\$100 range, we use the value 110 to represent in order to have the same intervals), and combine all values of the variable “employment_status” into two categories – employed and unemployed.

```{r q2_modified, echo=TRUE}
# modify "total_spend" column
q2_subset_naremoved$total_spend[q2_subset_naremoved$total_spend == '<$20'] = 10
q2_subset_naremoved$total_spend[q2_subset_naremoved$total_spend == '$20-$40'] = 30
q2_subset_naremoved$total_spend[q2_subset_naremoved$total_spend == '$40-$60'] = 50
q2_subset_naremoved$total_spend[q2_subset_naremoved$total_spend == '$60-$80'] = 70
q2_subset_naremoved$total_spend[q2_subset_naremoved$total_spend == '$80-$100'] = 90
q2_subset_naremoved$total_spend[q2_subset_naremoved$total_spend == '>$100'] = 110
q2_subset_naremoved$total_spend = as.numeric(q2_subset_naremoved$total_spend)

# modify "employment_status" column
q2_subset_naremoved$employment_status[q2_subset_naremoved$employment_status == 'Employed full-time'] = 'Employed'
q2_subset_naremoved$employment_status[q2_subset_naremoved$employment_status == 'Employed part-time'] = 'Employed'
q2_subset_naremoved$employment_status[q2_subset_naremoved$employment_status == 'Unemployed'] = 'Unemployed'
q2_subset_naremoved$employment_status[q2_subset_naremoved$employment_status == 'Student'] = 'Unemployed'
q2_subset_naremoved$employment_status[q2_subset_naremoved$employment_status == 'Retired'] = 'Unemployed'
q2_subset_naremoved$employment_status[q2_subset_naremoved$employment_status == 'Homemaker'] = 'Unemployed'

# the clean dataset for Question 2
q2_clean = q2_subset_naremoved
```

Now after these steps are completed, we create the faceted count plot using the code below.

```{r q2_plot, echo=TRUE}
# modify the texts of x-ticks (age groups)
q2_clean$age[q2_clean$age == '<18 years old'] = '<18'
q2_clean$age[q2_clean$age == '18-24 years old'] = '18-24'
q2_clean$age[q2_clean$age == '25-34 years old'] = '25-34'
q2_clean$age[q2_clean$age == '35-44 years old'] = '35-44'
q2_clean$age[q2_clean$age == '45-54 years old'] = '45-54'
q2_clean$age[q2_clean$age == '55-64 years old'] = '55-64'
q2_clean$age[q2_clean$age == '>65 years old'] = '>65'

# create the faceted count plot
ticks_order = c("<18", "18-24", "25-34", "35-44", "45-54", "55-64", ">65")
q2_clean$caffeine_factor = factor(q2_clean$caffeine, levels=c("Decaf", "Half caff", "Full caffeine"))

ggplot(q2_clean, aes(x=factor(age, level=ticks_order), total_spend)) +
  geom_count(aes(color=employment_status)) +
  scale_size_area(n.breaks = 10) +
  scale_color_manual(values=c("Employed"="#58DB65", "Unemployed"="#D21919")) +  
  labs(x="Age group (years old)", y="Spending on coffee ($ per month)", title = "Relationship Among Variables", size="Counts", col="Employment Status") +
  facet_grid(employment_status ~ caffeine_factor) +
  theme(panel.grid=element_line(color="grey90"), 
        panel.background=element_rect(fill="white", color="black"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) 
```

From the above visualization, we can see that there are some areas in the full caffeine facet where most of the data points are clustered, which means that a remarkable amount of surveyees prefer full caffeine to decaf or half-caff regardless of employment status. This overwhelming preference for full-caffeine coffee reflects people’s main demand when purchasing coffee, which is energizing strength and refreshing minds. This could originate from the fact that people often purchase coffee when they are very tired but need to recharge.

Then when we observe the distribution of monthly spending on coffee, we can see that regardless of employment status, surveyees who prefer half caff or full caffeine mostly spend 30-50 dollars per month, and surveyees who prefer decaf mostly spend less than or equal to 50 dollars per month. This could reflect the fact that people who prefer full caffeine might purchase coffee with stably higher frequencies than those who prefer lower caffeine. It aligns with the fact that caffeine can be addictive, and people who are used to full caffeine coffee might be more dependent on drinking coffee than those who prefer lower caffeine.

Furthermore, among the surveyees, the number of the employed is greater than the number of the unemployed. For the employed surveyees, most fall into 25-44 years old categories. For the unemployed surveyees, the age distribution is a bit more evenly distributed compared to the employed, and the 18-34 years old categories contain slightly more surveyees. This age distribution with respect to employment status in the dataset also reflects how overall employed people age above 25 years old when people become responsible adults and pursue careers. And there are unemployed people across all age groups, which also aligns with the bad situation in the job markets nowadays. This part of the discussion is not related to coffee preference, but it is also an interesting discovery from this visualization.

## Conclusion

From this project, we explored the Great American Coffee Taste Test dataset. We implemented the violin plot and the faceted count plot to display hidden trends, and we further extracted insights from these visualizations. These insights can be useful for coffee companies or any individuals who are interested in how the coffee preferences and consumption patterns vary from different demographics.
